External Patch Group Prior Guided Internal Prior Learning for Real Image Denoising

### 2016-10-26 CVPR2017_Review_Unified.tex
add

Given a real noisy image, we extract the noisy PGs and corresponding mean vectors. Each mean substracted PG is defined as $\mathbf{\overline{Y}}\triangleq \{\mathbf{\overline{x}}_{m}\}_{m=1}^{M}$. Noted that, different from the external PGs, the mean vectors of the internal noisy PGs are saved for recovering. For adaptivity, we project the PG $\mathbf{\overline{Y}}$ into its most suitable Gaussian component (subspace) of the GMM learned on external PGs. The subspace most suitable for $\mathbf{\overline{Y}}$ is selected by firstly calculating the posterior probability of "$\mathbf{\overline{Y}}$ belonging to the $k$th Gaussian component":
\vspace{-2mm}
\begin{equation}\label{equ5}\vspace{-2mm}
P(k|\mathbf{\overline{Y}})=\frac{\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{k})}{\sum_{l=1}^{K}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{l})},
\end{equation}
and then choosing the component with the maximum A-posteriori (MAP) probability $\ln P(k|\mathbf{\overline{Y}})$.

@@

Now all internal noisy PGs have been assigned to their corresponding most suitable Gaussians or subspaces in $\{\mathcal{N}(\mathbf{0},\mathbf{\Sigma}_{k})\}_{k=1}^{K}$. For each subspace, we employ its external orthogonal dictionary $\mathbf{D}_{e}$ to guide the internal orthogonal dictionary learning. The learning is performed under the sparse coding framework (along with which the denoising of real noisy image is simultaneously done) as follows:
\vspace{-2mm}
\begin{equation}\label{equ6}\vspace{-2mm}
\begin{split}
&\min_{\mathbf{D}_{i}\in \mathbb{R}^{3p^2\times r},\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
The singular vectors capture the statistical structures of NSS variations in natural images, while the singular values in $\mathbf{S}$ represent the significance of these singular vectors. Fig.\ 4 shows the singular vectors for one Gaussian component.

Similar to the K-SVD \cite{ksvd}, we employ an alternating iterative framework to solve the optimization problem \ref{equ4}. In fact, we initialize the orthogonal dictionary as $\mathbf{D}^{(0)}$ and for $t=0, 1, ..., T-1$, alternatively do:

\textbf{Updating Sparse Coefficients}: given the initialization orthogonal dictioanry $\textbf{D}_{i}^{(t)}$, the sparce coefficients 
$\textbf{A}^{(t)}$ are obtained via solving
\begin{equation}\label{equ6}
\textbf{A}^{(t)}:=\argmin_{\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}^{(t)}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}.
\end{equation}
This problem has closed-form solution by $\mathbf{A}^{*}=T_{\lambda}(\mathbf{\hat{D}}^{T}\mathbf{Y})$, 
where $T_{\lambda}(\mathbf{A}) = \text{sgn}(\mathbf{A})\odot\max(\mathbf{A},\lambda)$ is a soft-thresholding function.

\textbf{Updating Orthogonal Dictionary}: given the sparse coefficients $\textbf{A}^{(0)}$, the sparce coefficients 
$\textbf{A}^{(t)}$ are obtained via solving
\begin{equation}\label{equ7} 
\begin{split}
\textbf{D}_{i}^{(t+1)}:=
&
\argmin_{\textbf{D}_{i}\in\mathbb{R}^{3p^2\times r}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}]\mathbf{A}^{(t)}\|_{F}^{2}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
Dividing the sparse coefficients $\mathbf{A}=[\mathbf{A}_{e}^{T}\ \mathbf{A}_{i}^{T}]^{T}$, where $\mathbf{A}_{e}$ and $\mathbf{A}_{i}$ denote the coefficients over external and internal dictionary $\mathbf{D}_{e}$ and $\mathbf{D}_{i}$. According to the Proposition 2.2 in \cite{bao2013fast}, the problem (\ref{equ6}) 
%\begin{equation}
%\begin{split}
%\min_{\textbf{D}_{i}\in\mathbb{R}^{3p^2\times r}}
%&\|\mathbf{Y}-\mathbf{D}_{e}\mathbf{A}_{e}+\mathbf{D}_{i}\mathbf{A}_{i}\|_{F}^{2}
%\\
%&
%\text{s.t.}
%\quad
%\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
%\end{split}
%\end{equation}
has a closed-form solution $\mathbf{D}_{i}^{*}=\mathbf{U}\mathbf{V}^{T}$, where $\mathbf{U}$ and $\mathbf{V}$ are
the orthogonal matrices obtained by the following SVD
\begin{equation}\label{equ7}
(\mathbf{I}-\mathbf{D}_{e}\mathbf{D}_{e}^{T})\mathbf{Y}\mathbf{A}_{i}^{T}
=
\mathbf{U}\Sigma\mathbf{V}^{T}
\end{equation}
With these solutions, the final obtained dictionary $\mathbf{D} = [\mathbf{D}_{e}\ \mathbf{D}_{i}]$ are orthogonal ictionary. This can be proved by the following equation
\begin{equation}\label{equ8}
\mathbf{D}^{T}\mathbf{D} = 
\left(\begin{array}{c}
\mathbf{D}_{e}^{T}
\\
\mathbf{D}_{i}^{T}
\end{array}\right)
(\mathbf{D}_{e}\ \mathbf{D}_{i})
=
\left(\begin{array}{cc}
\mathbf{D}_{e}^{T}\mathbf{D}_{e}\ \mathbf{D}_{e}^{T}\mathbf{D}_{i}
\\
\mathbf{D}_{i}^{T}\mathbf{D}_{e}\ \mathbf{D}_{i}^{T}\mathbf{D}_{i}
\end{array}\right)
=
\mathbf{I}
\end{equation}

\subsection{Discussions on Learning External Priors and Internal Priors}
Until now, we have divided the noisy PGs into multiple internal subspaces. Here we take a deep analysis on how the external NSS prior guide the subspace learning of internal PGs. The help are at least threefold. Firstly, through MAP in (\ref{equ3}), the external prior guides the noisy PGs to be clustered into the correct subspaces. If we cluster the noisy PGs in an automatical way, the subspaces we learned will be highly degraded by the signal dependent noise. Secondly, the guidance of external prior for internal clustering is more efficient than directly clustering the internal noisy PGs. It only needs to calculate the MAP probability via the equation (\ref{equ3}) while the internal clustering via GMM is time-consuming on EM algorithm \cite{em}. Thirdly, due to the correct guidance of external prior, the structual decomposition via SVD of each subspace is more adaptive. This will bring better denoising performance than the methods only using the external information.

>>>>

Given a real noisy image, we extract the noisy PGs and corresponding mean vectors. Each mean substracted PG is defined as $\mathbf{\overline{Y}}\triangleq \{\mathbf{\overline{x}}_{m}\}_{m=1}^{M}$. Noted that, different from the external PGs, the mean vectors of the internal noisy PGs are saved for recovering. For adaptivity, we project the PG $\mathbf{\overline{Y}}$ into its most suitable Gaussian component (subspace) of the GMM learned on external PGs. The subspace most suitable for $\mathbf{\overline{Y}}$ is selected by firstly calculating the posterior probability of "$\mathbf{\overline{Y}}$ belonging to the $k$th Gaussian component":
\vspace{-2mm}
\begin{equation}\label{equ5}\vspace{-1mm}
P(k|\mathbf{\overline{Y}})=\frac{\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{k})}{\sum_{l=1}^{K}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{l})},
\end{equation}
and then choosing the component with the maximum A-posteriori (MAP) probability $\ln P(k|\mathbf{\overline{Y}})$.

@@

Now we have assigned all internal noisy PGs to their corresponding most suitable Gaussians or subspaces in $\{\mathcal{N}(\mathbf{0},\mathbf{\Sigma}_{k})\}_{k=1}^{K}$. For each subspace, we consider to use the external orthogonal dictionary to guide the learning of an orthogonal dictionary $\mathbf{D}:=[\mathbf{D}_{e}\ \mathbf{D}_{i}]\in \mathbb{R}^{3p^2\times 3p^2}$. This dictionary has two parts: the external part $\mathbf{D}_{e}\in\mathbb{R}^{3p^2\times (3p^2-r)}$ is consisted of the first ($3p^2-r$) columns of the singular vector matrix $\mathbf{U}_{e}$ obtained from the external prior by Equ. (\ref{equ3}), and the internal part $\mathbf{D}_{i}$ is consisted of adaptive dictionary atoms learned from the internal noisy PGs. The learning is performed under the sparse coding framework (along with which the denoising of real noisy image is simultaneously done) as follows:
\vspace{-2mm}
\begin{equation}\label{equ6}\vspace{-2mm}
\begin{split}
&\min_{\mathbf{D}_{i}\in \mathbb{R}^{3p^2\times r},\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
Similar to the K-SVD \cite{ksvd}, we employ an alternating iterative framework to solve the optimization problem \ref{equ4}. In fact, we initialize the orthogonal dictionary as $\mathbf{D}^{(0)}$ and for $t=0, 1, ..., T-1$, alternatively do:

\textbf{Updating Sparse Coefficients}: given the orthogonal dictioanry $\textbf{D}_{i}^{(t)}$, we update the sparce coefficients via solving
\vspace{-2mm}
\begin{equation}\label{equ7}\vspace{-2mm}
\textbf{A}^{(t)}:=\argmin_{\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}^{(t)}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}.
\end{equation}
This problem has closed-form solution by $\mathbf{A}^{*}=T_{\lambda}(\mathbf{\hat{D}}^{T}\mathbf{Y})$, 
where $T_{\lambda}(\mathbf{A}) = \text{sgn}(\mathbf{A})\odot\max(\mathbf{A},\lambda)$ is a soft-thresholding function.

\textbf{Updating Orthogonal Dictionary}: given the sparse coefficients $\textbf{A}^{(t)}$, we update the orthogonal dictionary via solving
\vspace{-2mm}
\begin{equation}\label{equ8} \vspace{-2mm}
\begin{split}
\textbf{D}_{i}^{(t+1)}:=
&
\argmin_{\textbf{D}_{i}\in\mathbb{R}^{3p^2\times r}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}]\mathbf{A}^{(t)}\|_{F}^{2}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
The sparse coefficients $\mathbf{A}=[\mathbf{A}_{e}^{T}\ \mathbf{A}_{i}^{T}]^{T}$ also has two parts: the external part $\mathbf{A}_{e}$ and the internal part $\mathbf{A}_{i}$ denote the coefficients over external dictionary $\mathbf{D}_{e}$ and internal dictionary $\mathbf{D}_{i}$, respectively. According to the Proposition 2.2 in \cite{bao2013fast}, the problem (\ref{equ6}) 
%\begin{equation}
%\begin{split}
%\min_{\textbf{D}_{i}\in\mathbb{R}^{3p^2\times r}}
%&\|\mathbf{Y}-\mathbf{D}_{e}\mathbf{A}_{e}+\mathbf{D}_{i}\mathbf{A}_{i}\|_{F}^{2}
%\\
%&
%\text{s.t.}
%\quad
%\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
%\end{split}
%\end{equation}
has a closed-form solution $\mathbf{D}_{i}^{*}=\mathbf{U}\mathbf{V}^{T}$, where $\mathbf{U}$ and $\mathbf{V}$ are
the orthogonal matrices obtained by the following SVD
\vspace{-2mm}
\begin{equation}\label{equ9}\vspace{-2mm}
(\mathbf{I}-\mathbf{D}_{e}\mathbf{D}_{e}^{T})\mathbf{Y}\mathbf{A}_{i}^{T}
=
\mathbf{U}\Sigma\mathbf{V}^{T}
\end{equation}
With these solutions, the final obtained dictionary $\mathbf{D} = [\mathbf{D}_{e}\ \mathbf{D}_{i}]$ are orthogonal ictionary. This can be proved by the following equation
\vspace{-3mm}
\begin{equation}\label{equ10}\vspace{-3mm}
\mathbf{D}^{T}\mathbf{D} = 
\left[\begin{array}{c}
\mathbf{D}_{e}^{T}
\\
\mathbf{D}_{i}^{T}
\end{array}\right]
[\mathbf{D}_{e}\ \mathbf{D}_{i}]
=
\left[\begin{array}{cc}
\mathbf{D}_{e}^{T}\mathbf{D}_{e}\ \mathbf{D}_{e}^{T}\mathbf{D}_{i}
\\
\mathbf{D}_{i}^{T}\mathbf{D}_{e}\ \mathbf{D}_{i}^{T}\mathbf{D}_{i}
\end{array}\right]
=
\mathbf{I}
\end{equation}

\subsection{Discussions}
Until now, we have divided the noisy PGs into multiple internal subspaces. Here we take a deep analysis on how the external NSS prior guide the subspace learning of internal PGs. The help are at least threefold. Firstly, through MAP in (\ref{equ3}), the external prior guides the noisy PGs to be clustered into the correct subspaces. If we cluster the noisy PGs in an automatical way, the subspaces we learned will be highly degraded by the signal dependent noise. Secondly, the guidance of external prior for internal clustering is more efficient than directly clustering the internal noisy PGs. It only needs to calculate the MAP probability via the equation (\ref{equ3}) while the internal clustering via GMM is time-consuming on EM algorithm \cite{em}. Thirdly, due to the correct guidance of external prior, the structual decomposition via SVD of each subspace is more adaptive. This will bring better denoising performance than the methods only using the external information.
